---
output:
  html_document:
    theme: flatly
---
<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Tutorial for the JRC's eSpecies API}
-->

```{r setup, echo=FALSE}
options(width=160)
```


### Basic information and statistics

DOPA provides an array of information on species and protected area (PA) 
occurrence within the borders (or in some cases the Exclusive Economic Zone 
boundary)  of a given country. `country_list()` can be used to list all the 
countries available in the DOPA database:

```{r country-list, message=FALSE}
library(rdopa)
# We'll also be using package dplyr for data manipulation
library(dplyr)

cl <- country_list()
head(cl)
```

### Species 

There are also few convience function to retrieve species information on 
conuntry-level. To list species included in a particular 
[IUCN status category](http://www.iucnredlist.org/technical-documents/categories-and-criteria)
within a given country, use function `country_species_list()`. For example, if 
you're interested in all (globally) threatened species (i.e. those in IUCN 
categories CR, EN, or VU) that occur in Finland, do the following:

```{r species-list-1, message=FALSE}
threatened.fin <- country_species_list('Finland', rlstatus=c("CR", "EN", "VU"))
# Select only part of the columns (using dplyr)
select(threatened.fin, iucn_species_id, taxon, class, status, assessed, commonname)
```

Function `country_species_count()` returns the total number of species within
a given country that belong to a specific status categories. For example, to 
get the number of threatened species in Brazil:

```{r species-count-1, message=FALSE}
country_species_count('Brazil', rlstatus=c("CR", "EN", "VU"))
```

This - and breakdown into individual categories - can of course be achieved by
working with the complete country-specific species list:

```{r species-list-2, message=FALSE}
threatened.bra <- country_species_list('Brazil', rlstatus=c("CR", "EN", "VU"))
# Total number
nrow(threatened.bra)
# Per category (using dplyr)
threatened.bra %>%
  group_by(status) %>%
  summarise(count = n()) 
```

### Protected areas

Protected area (PA) statistics are available per country according to 
[IUCN PA categories](http://www.iucn.org/about/work/programmes/gpap_home/gpap_quality/gpap_pacategories/):

```{r country-stats, message=FALSE}
country_stats("Sweden")
```

Within a given country, statistics for individual PAs can be retrieved using
`pa_country_stats()`. Let's find out all the PAs within Uganda:

```{r country-pa_stats, message=FALSE}
uganda_pa_stats <- pa_country_stats("Uganda")
nrow(uganda_pa_stats)
```

So Uganda has 28 individual PAs in the WDPA. 
