<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{Tutorial for the JRC's eSpecies API}
-->

```{r setup, echo=FALSE}
options(width=160)
```

## Country-level information

### Basic information and statistics

DOPA provides an array of information on species and protected area (PA) 
occurrence within the borders (or in some cases the Exclusive Economic Zone 
boundary)  of a given country. `country_list()` can be used to list all the 
countries available in the DOPA database:

```{r country-list, message=FALSE}
library(rdopa)
# We'll also be using package dplyr for data manipulation
library(dplyr)

cl <- country_list()
head(cl)
```

### Species 

There are also few convience function to retrieve species information on 
conuntry-level. To list species included in a particular 
[IUCN status category](http://www.iucnredlist.org/technical-documents/categories-and-criteria)
within a given country, use function `country_species_list()`. For example, if 
you're interested in all (globally) threatened species (i.e. those in IUCN 
categories CR, EN, or VU) that occur in Finland, do the following:

```{r species-list-1, message=FALSE}
threatened.fin <- country_species_list('Finland', rlstatus=c("CR", "EN", "VU"))
select(threatened.fin, iucn_species_id, taxon, class, status, assessed, commonname)
```
